Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-02-10 Paper/Paper.tex
updates results

\subsection{Exploring the Model}\label{ExplModel}
In Figure \ref{fig:ModelIntro}A a schematic description of the model is shown.
In the right half of Figure \ref{fig:ModelIntro}A one can see the cell cycle choices that each cell will make each cell cycle and how the various components of the model will influence these probabilities.
The inset on the left gives an overview of how the immune system works.
The dashed arrows indicate that the action is probabilistic rather than deterministic.
In other words, the cytotoxic cells will randomly clear some subset of mutated cells and TGF-$\beta$ increases the probability of cell transitioning or staying in a mesenchymal state.

For a typical {\it in silico} patient, see Figure \ref{fig:ModelIntro}B.
The inflammation cycling scheme shown at the top of Figure \ref{fig:ModelIntro}B is repeated throughout.
We will return to this shortly.
It is at cell cycle number 541 the proportion of mutated cells reaches 50\% and so the patient has a Time to Cancer of 541, or 405.75 days. 
At the same time as mutated cells are taking over, most cells are transitioning to a mesenchymal phenotype.
After the Time to Cancer, the mutated cells continue their rapid growth and soon make up 100\% of the cells.
Meanwhile, all the cells become mesenchymal for a short duration before most transition back to an epithelial state.

Considering the immune populations, notice that the NK population basically stays constant while the lymphocyte populations quickly dwarf the NK population once cells begin to mutate.
These lymphocyte populations appear to exhibit oscillatory behavior, however this is not due to intrinsic dynamics but rather due to the inflammation scheme that this patient is undergoing: 
the patient alternates between 30 cell cycles of high inflammation and 60 cell cycles of low inflammation.
In each of these periods periods, the lymphocyte populations increase or decrease rapidly in accordance with the inflammation.

Finally, considering a cohort of $N=500$ patients, we can generate a survival curve for these patients tracking how many have survived cancer-free up to a give day.
See Figure \ref{fig:ModelIntro}C for such a survival curve.
In this curve, we see that all patients survive for some minimum amount of time before a time period a nearly constant cancer onset rate roughly between $T = 100$ and $T = 300$.

\subsection{Sensitivity Analysis}\label{SensAnalysis}
To assess the sensitivity of the model to the 31 model parameters, a Morris OAT algorithm was implemented.
The algorithm works by running the model many successive times with exactly one parameter changing between each run\cite{morris1991factorial}\cite{sohier2014improvement}.
The parameters were all given Gaussian distributions centered at their initial value and a large standard deviation.
Those that needed bounds were truncated.
The results of the Morris OAT can be found in Figure \ref{fig:ModelIntro}D.
Clearly, a certain subset of parameters have an outsized effect on the model while others do relatively little.
The two most influential according to this analysis are related to Treg cells: their death rate and recruitment.
This makes sense as Treg cells both suppress cytotoxic effects of other immune cells and also secrete significant quantities of TGF-$\beta$ which drives EMT.
Thus, Treg cells are intimately tied to all three aspects of the model and so their presence or lack thereof significantly affects the system.
However, we want to better isolate the effects of EMT on these dynamics, so parameters such as MIE and MGA are of interest.
In addition, inflammation parameters dictating the cycling scheme are worthy of study because of their prominence in Guo's work.
What we can see from this plot is that the model also has some high sensitivity to these parameters.
In terms of Treg cells, their secretion of TGF-$\beta$ is highly significant and so we will also look at that parameter.

>>>>

\subsection{A multiscale agent-based model of EMT-immune-tissue cell interactions and tumorigenesis}\label{ExplModel}
We begin by studying the model under baseline conditions to assess how its various components interact and effect the overal probability of tumorigenesis (Time to Cancer). In Figure \ref{fig:ModelIntro}A a schematic description of the model is shown. Within a single cell cycle, different choices are available to a cell and these are influenced by pathway mutations or EMT. For example, if a cell undergoes EMT, the probability that is will proliferate is reduced. Also shown are the different means by which the immune system acts on the tissue cells (\ref{fig:ModelIntro}A Inset). \tcr{TGF-$\beta$ is [describe briefly action of each immune cell type]}
In other words, the cytotoxic cells will randomly clear some subset of mutated cells and TGF-$\beta$ increases the probability of cell transitioning or staying in a mesenchymal state.

For a typical {\it in silico} patient (see Figure \ref{fig:ModelIntro}B) \tcr{[What happens??]}
The inflammation cycling scheme shown at the top of Figure \ref{fig:ModelIntro}B is repeated throughout.
We will return to this shortly.

\tcr{[What is?]} It is at cell cycle number 541 the proportion of mutated cells reaches 50\% and so the patient has a Time to Cancer of 541, or 405.75 days. 
At the same time as mutated cells are taking over, many cells transition to a mesenchymal phenotype.
After the Time to Cancer, the mutated cells continue their rapid growth and soon make up 100\% of the cells.
Meanwhile, all the cells become mesenchymal for a short duration before most transition back to an epithelial state.

Considering the immune populations, we see that the NK population is approximately constant, while the lymphocyte populations grow quickly and dwarf the NK population following the accumulation of mutations.
The lymphocyte populations exhibit oscillatory behavior, however note that this is not due to intrinsic dynamics but rather due to the inflammation scheme that this patient is undergoing: 
the patient alternates between 30 cell cycles of high inflammation and 60 cell cycles of low inflammation.
In each of these periods periods, the lymphocyte populations increase or decrease rapidly in accordance with the inflammation.

We next consider a cohort of $N=500$ patients, and simulate the survival curve for these patients (Figure \ref{fig:ModelIntro}C). We see that all patients survive for approximately 100 cell cycles (75 days), before a time period where the cancer onset rate is roughly constant, between the approximate range of ($T = 100, T = 300$).

\subsection{Identification of key model parameters via Morris global sensitivity analysis}\label{SensAnalysis}
To assess the sensitivity responses produced by the model with respect to the 31 model parameters, a Morris one-step-at-a-time (OAT) algorithm was implemented. \tcr{add paragraph to methods on this.}
The algorithm works by running the model many successive times, changing exactly one parameter each run \cite{morris1991factorial, sohier2014improvement}. Parameters are sampled from Gaussian distributions centered on their initial \tcr{(which is determined how?)} value and a large \tcr{(specify)} standard deviation. Where necessary, these distributions were truncated accordingly.
The results of Morris OAT sensitivity analysis are shown in Figure \ref{fig:ModelIntro}D. A subset of parameters demonstrate much higher levels of sensitivity than others.
The two most influential according to this analysis are the death rate and the recruitment rate of Treg cells, this is most likely due to the dual roles Treg cells play in both suppressing the cytotoxic effects of other immune cells and secreting TGF-$\beta$, which drives EMT. This ties Treg cells to all three components of the model. Since below we seek to separate the effects of different model components, we do not choose the parameters influencing Treg cells for further analysis below. 
We want to better isolate the effects of EMT on these dynamics, so parameters such as MIE and MGA are of interest.
In addition, inflammation parameters dictating the cycling scheme are important to study, as shown by Guo et al \cite{}: we find that there are also high sensitivities in this model to these parameters.
In terms of Treg cells, their secretion of TGF-$\beta$ is highly significant and we will also study it further.
